### 2014-12-04 paper/paper.tex
Added LaTeX paper draft

\documentclass[fontsize=11pt]{scrartcl}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}				% special symbols
\usepackage{fancyhdr}					% headers and footers
\usepackage{verbatim}					% for comments
\usepackage{graphicx}					% images
\usepackage[font=footnotesize]{caption}	% smaller image label
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor} 					% highlighting
\usepackage{hyperref}				    % for links (text or internet)
\usepackage{amsmath}	    			% maths
\usepackage{relsize}

\usepackage[hyperref=true,style=authoryear,backend=biber]{biblatex}
\usepackage{filecontents}
\usepackage{setspace}
\usepackage{fancyhdr}

%\linespread{2}

% \pagestyle{headings}

\addbibresource{bibliography.bib}

\title{Real Time Fluid Simulation \\ \smaller using Smoothed-Particle Hydrodynamics and OpenGL}

\subtitle{Computer Graphics CS 488}

\author{Michael Berg \\ Matthias Untergassmair}

\pagestyle{fancy}
\rhead{\thepage}

\begin{document}

\maketitle

\abstract{
	This paper talks about the implementation of a real time Fluid simulation displayed in OpenGL \dots blaaaaaaaaaaaaa
}

\clearpage

\section{Introduction}
%
Hello World! This is how you can cite a paper \cite{realtimerendering} Let's go to a new line \\
Let's go to a new paragraph \par\medskip
Let's write a formula
\begin{align*}
	\sum_{i=0}^N \int_{-\infty}^\infty e^{-\frac{1}{2}x^2} \; dx
\end{align*}
%
It's not that hard ;)

\section{Mathematical Background}

\section*{Appendix: Code}

\clearpage
\nocite{*}
\printbibliography

\end{document}

### 2014-12-04 Simulation.cpp
Added Article to folder

@@ -49,7 +49,7 @@ class Simulation : public Renderer {
 
 public:
 
-	static const unsigned N = 800;
+	static const unsigned N = 2000;
 	unsigned M = 0;
 		
 	ResourceHandler rh; 

### 2014-12-05 Simulation.cpp
minor change in shader

@@ -90,8 +90,8 @@ public:
 		for(int i=0; i<M; ++i) {
 			MeshData md;
 			// addCube(md,fluidsimulation.getRadius(i),vec3(0,0,0));
-			// addRect(md,10.f,30.f,50.f,vec3(0,0,0));
-			addSphere(md,fluidsimulation.getRadius(i),8,8);
+			addRect(md,4.f,4.f,100.f,vec3(0,0,0));
+			// addSphere(md,fluidsimulation.getRadius(i),8,8);
 			mb[i].init(md,posLoc,normalLoc,-1,colLoc);
 		}
 
@@ -200,7 +200,7 @@ public:
 				glUniformMatrix4fv(program.uniform("view"), 1, 0, ptr(camera.view));
 				glUniformMatrix4fv(program.uniform("proj"), 1, 0, ptr(camera.projection));
 
-				glUniform3f(program.uniform("velocity"), std::abs(velocity[0])/100,velocity[1]/100,0);
+				glUniform3f(program.uniform("velocity"), std::abs(velocity[0])/100,velocity[1]/100,.25);
 
 				mb[i].draw();
 			} program.unbind();

### 2014-12-07 paper/paper.tex
Velocity Verlet Implemented, but still has bugs + Added velocity verlet scheme to report

@@
 \usepackage{fancyhdr}
+\usepackage[ruled,vlined]{algorithm2e}
 
@@
 \section{Mathematical Background}
+%
+\begin{align}
+	a_i^n = \frac{F_i^n}{m_i} = \dots
+	\label{math:acceleration}
+\end{align}
+
+\section{The Algorithm}
+%
+In the following, we denote the position for the particle $i$ at time $t$ as $x_i^t$, its velocity as $v_i^t$ and its acceleration as $a_i^t$.
+We omit the vector notation ($\boldsymbol{x},\boldsymbol{v},\boldsymbol{a}$) for these quantities, since the following equations are valid for the vectors as well as for each component individually.
+\par\medskip
+%
+As suggested in \cite{generalizedwallboundary}, we use the \emph{Velocity-Verlet} time stepping scheme as follows:
+%
+\begin{algorithm}
+	\caption{Single Timestep with Velocity Verlet Algorithm}
+	\KwData{$x_i^t$, $v_i^{t - \frac{\Delta t}{2}}$, $a_i^t$, $\Delta t$}
+	\KwResult{$x_i^{t + \Delta t}$, $v_i^{t + \frac{\Delta t}{2}}$, $a_i^{t + \Delta t}$}	
+	$v_i^{t+\frac{\Delta t}{2}} = v_i^{t - \frac{\Delta t}{2}} + \Delta t a_i^t$ \;
+	$x_i^{t+\Delta t} = x_i^t + \Delta t v_i^{t+\frac{\Delta t}{2}}$ \;	
+	$a_i^{t + \Delta t} = a_i^{t + \Delta t} (x_i^{t + \Delta t}, m_i)$ from equation \ref{math:acceleration} \;
+\end{algorithm}
 

### 2014-12-07 Simulation.cpp

@@ -49,7 +49,7 @@ class Simulation : public Renderer {
 
 public:
 
-	static const unsigned N = 2000;
+	static const unsigned N = 32;
 	unsigned M = 0;
 		
 	ResourceHandler rh; 
@@ -90,8 +90,8 @@ public:
 		for(int i=0; i<M; ++i) {
 			MeshData md;
 			// addCube(md,fluidsimulation.getRadius(i),vec3(0,0,0));
-			addRect(md,4.f,4.f,100.f,vec3(0,0,0));
-			// addSphere(md,fluidsimulation.getRadius(i),8,8);
+			// addRect(md,4.f,4.f,100.f,vec3(0,0,0));
+			addSphere(md,fluidsimulation.getRadius(i),8,8);
 			mb[i].init(md,posLoc,normalLoc,-1,colLoc);
 		}
 
@@ -129,7 +129,7 @@ public:
 		// PROPAGATE MODEL
 		///////////////////////////////////////////////////////////
 
-		if(stepCounter < 50000) {
+		if(stepCounter < 5000) {
 			++stepCounter;
 			fluidsimulation.timestep(.05); // Propagate fluidsimulation in time
 			std::cout << fluidsimulation; // Output current status of Fluid particles

### 2014-12-07 sphModel.cpp

@@ -14,6 +14,9 @@ SPH::SPH(unsigned N)
 		_v1(new float[_nTotal]),
 		_v2(new float[_nTotal]),
 		_v3(new float[_nTotal]),
+		_a1(new float[_nTotal]),
+		_a2(new float[_nTotal]),
+		_a3(new float[_nTotal]),
 		_m(new float[_nTotal]),
 		_r(new float[_nTotal]),
 		_x1MinWall(-100),
@@ -26,7 +29,6 @@ SPH::SPH(unsigned N)
 		_x2MaxBox(+100),
 		_g(0),
 		_damping(.8),
-		_ready(true),
 		_T(0.0),
 		_tStep(0),
 		_dt(0),
@@ -46,20 +48,24 @@ SPH::SPH(unsigned N)
 	for(unsigned i=0; i<_nParticles; ++i) {
 		
 		// Position
-		_x1[i] = (i%2 ? -50 : 50); // 10*float(i)/_nParticles;
+		_x1[i] = (i%2 ? -50 : 50) + .1*dist(e2); // 10*float(i)/_nParticles;
 		_x2[i] = 0;
 		_x3[i] = 0;
 		
-		// Velocities (sampled from random normal distribution)
-		_v1[i] = dist(e2);
-		_v2[i] = dist(e2);
-		_v3[i] = 0;
-
 		// Masses (assume all particles have the same mass)
 		_m[i] = 1e-6 * (1+i%3);
 
 		// Radius / Support of particles
 		_r[i] = 1+i%3;
+
+		// Compute Forces acting on particles based on positions
+		updateForces();
+		
+
+		// Velocities (sampled from random normal distribution)
+		_v1[i] = 0; // 0.1*dist(e2);
+		_v2[i] = 0; // dist(e2);
+		_v3[i] = 0;
 
 		// _v1[i] = 0; // TODO: remove, v_x = 0 only for debugging
 		// _v2[i] = 40.f; // TODO: remove, v_y = 40 only for debugging
@@ -99,7 +105,7 @@ SPH::SPH(unsigned N)
 		
 		_x3[i] = 0;
 		
-		// Velocities (sampled from random normal distribution)
+		// Velocities = 0 in boundary
 		_v1[i] = 0;
 		_v2[i] = 0;
 		_v3[i] = 0;
@@ -148,7 +154,7 @@ SPH::SPH(unsigned N)
 		}
 		
 		
-		// Velocities (sampled from random normal distribution)
[…]

### 2014-12-07 sphModel.hpp

@@ -1,6 +1,7 @@
 #ifndef SPH_HPP
 #define SPH_HPP
 
+#define _USE_MATH_DEFINES // make M_PI available
 
 #include <iostream>
 #include <stdio.h>
@@ -19,9 +20,12 @@ class SPH {
 		
 		// Destructor
 		~SPH(); 
-
+		
 		// Time Propagation of model
 		void timestep(float);
+		
+		// Update forces on particles based on SPH
+		void updateForces();
 
 		// Overloading Output Operator
 		friend ostream& operator<<(ostream&, const SPH&);
@@ -29,9 +33,6 @@ class SPH {
 		// Applying elastic boundary conditions
 		void applyBoundary();
 
-		// Return status of model - Ready for rendering?
-		bool isReady() const;
-		
 		// Get Radius of Particle i
 		unsigned getTotalParticles() const;
 
@@ -80,13 +81,16 @@ class SPH {
 		unsigned _nGhostObject; // Number of ghost particles in the object
 		unsigned _nTotal; // Total number of particles
 
-		// Array of particle coordinates & velocities
+		// Array of particle coordinates & velocities & accelerations
 		float* _x1;
 		float* _x2;
 		float* _x3;
 		float* _v1;
 		float* _v2;
 		float* _v3;
+		float* _a1;
+		float* _a2;
+		float* _a3;
 
 		// Array of particle masses
 		float* _m;
@@ -116,9 +120,6 @@ class SPH {
 
 		// Damping factor for elastic bounding on walls
 		float _damping;
-
-		// Boolean flag that signalizes whether the model is computing or idle
-		bool _ready;
 
 		// Total time
 		float _T;

### 2014-12-10 paper/paper.tex
Implemented time stepping + added 3d force + added code to paper appendix + added some articles

@@
 \usepackage[ruled,vlined]{algorithm2e}
+
+\usepackage{listings}
+
+\lstset{language=C++,
+    basicstyle=\footnotesize\ttfamily,
+    keywordstyle=\bfseries\color{blue},%
+    showstringspaces=false,
+    breaklines=true,%
+    morekeywords={include, printf},
+	backgroundcolor=\color{lightgray},
+    identifierstyle=\color{black},%
+    stringstyle=\color{DarkOrchid},
+    commentstyle=\color{Green},%
+    numbers=left,
+    numberstyle={\tiny \color{black}},
+    numbersep=9pt,
+    emph=[1]{for,if,else,switch,case,while,return},emphstyle=[1]\color{red},
+	stepnumber=5,
+	tabsize=2
+}
 
@@
 
+\clearpage
+
 \section*{Appendix: Code}
+
+\lstinputlisting[caption=sphModel.hpp]{../sphModel.hpp}
+\lstinputlisting[caption=sphModel.cpp]{../sphModel.cpp}
+\lstinputlisting[caption=Simulation.cpp]{../Simulation.cpp}
+\lstinputlisting[caption=simulation.vsh]{../resources/simulation.vsh}
+\lstinputlisting[caption=simulation.fsh]{../resources/simulation.fsh}
 

### 2014-12-10 Simulation.cpp

@@ -21,7 +21,7 @@
 	#include "Aluminum/Behavior.hpp"
 	#include "Aluminum/ResourceHandler.hpp"
 	#include "Aluminum/Texture.hpp"
-	#include "Aluminum/Renderer.hpp"
+	#include "Aluminum/RendererLinux.hpp"
 #elif defined _WIN32 || defined _WIN64
 	std::string platform = "WINDOWS";
 #else
@@ -44,12 +44,16 @@ float pi = glm::pi<float>();
 using namespace aluminum;
 
 
+// TODO: improve performance by only adding one single sphere instead of N spheres (just use different model matrices).
+// TODO: make liquid flow in from top
+// TODO: pass in only points to shader and use geometry shader to create 3d particles
+// TODO: see 3.5.1: flowing water and particle effects, stream output
 
-class Simulation : public Renderer {
+class Simulation : public RendererLinux {
 
 public:
 
-	static const unsigned N = 32;
+	static const unsigned N = 0; // 40;
 	unsigned M = 0;
 		
 	ResourceHandler rh; 
@@ -91,7 +95,7 @@ public:
 			MeshData md;
 			// addCube(md,fluidsimulation.getRadius(i),vec3(0,0,0));
 			// addRect(md,4.f,4.f,100.f,vec3(0,0,0));
-			addSphere(md,fluidsimulation.getRadius(i),8,8);
+			addSphere(md,5*fluidsimulation.getRadius(i),8,8);
 			mb[i].init(md,posLoc,normalLoc,-1,colLoc);
 		}
 
@@ -129,7 +133,7 @@ public:
 		// PROPAGATE MODEL
 		///////////////////////////////////////////////////////////
 
-		if(stepCounter < 5000) {
+		if(stepCounter < 0 /*5000*/) {
 			++stepCounter;
 			fluidsimulation.timestep(.05); // Propagate fluidsimulation in time
 			std::cout << fluidsimulation; // Output current status of Fluid particles
@@ -187,7 +191,7 @@ public:
 				fluidsimulation.getPosition(i,position);
 				fluidsimulation.getVelocity(i,velocity);
 
-				model = glm::translate(model,vec3(position[0],position[1],0));
+				model = glm::translate(model,vec3(position[0],position[1],position[2]));
 
 				// For Rotation of Cubes
 				/*
[…]

### 2014-12-10 sphModel.cpp

@@ -6,7 +6,7 @@
 SPH::SPH(unsigned N)
 	:	_nParticles(N),
 		_nGhostWall(100),
-		_nGhostObject(100),
+		_nGhostObject(0),
 		_nTotal(_nParticles+_nGhostWall+_nGhostObject),
 		_x1(new float[_nTotal]),
 		_x2(new float[_nTotal]),
@@ -19,14 +19,14 @@ SPH::SPH(unsigned N)
 		_a3(new float[_nTotal]),
 		_m(new float[_nTotal]),
 		_r(new float[_nTotal]),
-		_x1MinWall(-100),
-		_x1MaxWall(+100),
 		_x2MinWall(-100),
 		_x2MaxWall(+100),
-		_x1MinBox(-40),
-		_x1MaxBox(+40),
-		_x2MinBox(-100),
-		_x2MaxBox(+100),
+		_x3MinWall(-100),
+		_x3MaxWall(+100),
+		_x2MinBox(-40),
+		_x2MaxBox(+40),
+		_x3MinBox(-100),
+		_x3MaxBox(+100),
 		_g(0),
 		_damping(.8),
 		_T(0.0),
@@ -48,23 +48,22 @@ SPH::SPH(unsigned N)
 	for(unsigned i=0; i<_nParticles; ++i) {
 		
 		// Position
-		_x1[i] = (i%2 ? -50 : 50) + .1*dist(e2); // 10*float(i)/_nParticles;
-		_x2[i] = 0;
-		_x3[i] = 0;
+		_x1[i] = 0;
+		_x2[i] = (i%2 ? -50 : 50) + .1*dist(e2);
+		_x3[i] = fmod(rand(),200)-100;
 		
 		// Masses (assume all particles have the same mass)
-		_m[i] = 1e-6 * (1+i%3);
+		_m[i] = 1; // 1e-6 * (1+i%3);
 
 		// Radius / Support of particles
-		_r[i] = 1+i%3;
+		_r[i] = 1; // 1+i%3;
 
 		// Compute Forces acting on particles based on positions
 		updateForces();
-		
 
 		// Velocities (sampled from random normal distribution)
-		_v1[i] = 0; // 0.1*dist(e2);
-		_v2[i] = 0; // dist(e2);
+		_v1[i] = dist(e2);
+		_v2[i] = dist(e2);
 		_v3[i] = 0;
 
 		// _v1[i] = 0; // TODO: remove, v_x = 0 only for debugging
[…]

### 2014-12-10 sphModel.hpp

@@ -14,6 +14,8 @@ using std::ostream;
 class SPH {
 
 	public:
+
+		static const unsigned _ghostDepth = 3;
 		
 		// Constructor
 		SPH(unsigned); 
@@ -42,9 +44,10 @@ class SPH {
 				std::cout << "ERROR: Invalid index";
 				return;
 			}
-			x[0] = _x1[index];
-			x[1] = _x2[index];
-			x[2] = _x3[index];
+			// Take into account switching of axes for OpenGL
+			x[0] = _x2[index];
+			x[1] = _x3[index];
+			x[2] = _x1[index];
 		}
 		
 		// Write velocity to the 3-array v
@@ -53,9 +56,10 @@ class SPH {
 				std::cout << "ERROR: Invalid index";
 				return;
 			}
-			v[0] = _v1[index];
-			v[1] = _v2[index];
-			v[2] = _v3[index];
+			// Take into account switching of axes for OpenGL
+			v[0] = _v2[index];
+			v[1] = _v3[index];
+			v[2] = _v1[index];
 		}
 
 		// Return Kinetic Energy
@@ -71,8 +75,8 @@ class SPH {
 		void setGravity(float);
 
 		// Functions for changing Box position
-		void moveBoxX(float);
-		void moveBoxY(float);
+		// void moveBoxX(float);
+		// void moveBoxY(float);
 
 	private:
 
@@ -99,20 +103,20 @@ class SPH {
 		float* _r;
 
 		// Wall Coordinates
-		float _x1MinWall;
-		float _x1MaxWall;
 		float _x2MinWall;
 		float _x2MaxWall;
+		float _x3MinWall;
+		float _x3MaxWall;
 
 		// Box Coordinates
-		float _x1MinBox;
-		float _x1MaxBox;
 		float _x2MinBox;
 		float _x2MaxBox;
+		float _x3MinBox;
+		float _x3MaxBox;
 
 		// Velocity component introduced by Box movement
-		float _v1Box;
 		float _v2Box;
+		float _v3Box;
 
 
 		// Gravity

### 2014-12-10 Simulation.cpp
README updated

@@ -167,7 +167,8 @@ public:
 		// Start displaying
     
 		glViewport(0, 0, width, height);
-		glClearColor(0.1,0.1,0.1,1.0);
+		// glClearColor(0.1,0.1,0.1,1.0);
+		glClearColor(1.0,1.0,1.0,1.0);
 		glClear(GL_COLOR_BUFFER_BIT | GL_DEPTH_BUFFER_BIT);	
 		
 		if (camera.isTransformed) {

### 2014-12-11 paper/paper.tex
Added intro paragraph and Astrophysics paper

@@
 \abstract{
-	This paper talks about the implementation of a real time Fluid simulation displayed in OpenGL \dots blaaaaaaaaaaaaa
+	This paper explores the field of Smooth Particle Hydrodynamics (SPH), starting at its 
+beginnings as a tool to simulate astrophysical phenomena and following its evolution and 
+implementation as a way to simulate fluids such as water. We will then give a brief 
+mathematical background and ensuing algorithm of our SPH simulation followed by a detailed 
+explanation of it.
 }
@@
 %
-Hello World! This is how you can cite a paper \cite{realtimerendering} Let's go to a new line \\
-Let's go to a new paragraph \par\medskip
-Let's write a formula
-\begin{align*}
-	\sum_{i=0}^N \int_{-\infty}^\infty e^{-\frac{1}{2}x^2} \; dx
-\end{align*}
+\hspace{6 pt} Smoothed Particle Hydrodynamics (SPH) successfully simulates fluids by breaking up a 
+fluid body into individual parts, or particles. These particles together form a particle 
+system that simulates various gravitational forces. Fluid movement is simulated in the 
+system by moving particles around any particle moved, simulating a rippling, wave effect.
+\\
+\hspace*{6 pt} The ripple effect is created by first calculating which particles are surrounding 
+a particle moved, and these surrounding particles are moved according to the movement 
+of the first particle moved. But SPH wasn't originally intended to simulate liquid 
+substances, but astrophysical phenomena.
 %
-It's not that hard ;)
+
+\section{History}
+%
+something
+%
 

### 2014-12-11 paper/paper.tex
Added history section

@@
 %
-something
+\hspace{6 pt} Given all the different applications for Smoothed Particle Hydrodynamics (SPH),
+it was first used to simulate interstellar phenomena. Conceived in 1977 by Gingold and 
+Monaghan was an improvement to the Standard Finite Difference Method, which until their 
+breakthrough, was the method to use to simulate astrophysical phenomena. They improved 
+on this method by making "use of Lagrangian description of fluid flow which automatically 
+focuses attention on fluid elements" \cite{sphastrophysics}. In this implementation, particles 
+"move according to the Newtonian equations with forces due to the pressure gradient and 
+other body forces: gravity, rotation and magnetic" \cite{sphastrophysics}.
+\\
+\hspace*{6 pt} 
 %

### 2014-12-12 paper/paper.tex
Finished off History section, updated and compiled bibliography
Put paper in requested format

@@
-\documentclass[fontsize=11pt]{scrartcl}
+\documentclass{sigchi}
 \usepackage[english]{babel}
@@
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor} 					% highlighting
-\usepackage{hyperref}				    % for links (text or internet)
 \usepackage{amsmath}	    			% maths
@@
 
-\usepackage[hyperref=true,style=authoryear,backend=biber]{biblatex}
+\usepackage[hyperref=true,style=sigchi,backend=biber,doititles]{biblatex}
 \usepackage{filecontents}
@@
 \usepackage[ruled,vlined]{algorithm2e}
+
+\usepackage{balance}  % to better equalize the last page
+\usepackage{times}    % comment if you want LaTeX's default font
+\usepackage{url}      % llt: nicely formatted URLs
 
@@
 
+% llt: Define a global style for URLs, rather that the default one
+\makeatletter
+\def\url@leostyle{%
+  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
+\makeatother
+\urlstyle{leo}
+
+% To make various LaTeX processors do the right thing with page size.
+\def\pprw{8.5in}
+\def\pprh{11in}
+\special{papersize=\pprw,\pprh}
+\setlength{\paperwidth}{\pprw}
+\setlength{\paperheight}{\pprh}
+\setlength{\pdfpagewidth}{\pprw}
+\setlength{\pdfpageheight}{\pprh}
+
+% Make sure hyperref comes last of your loaded packages, 
+% to give it a fighting chance of not being over-written, 
+% since its job is to redefine many LaTeX commands.
+\usepackage[pdftex]{hyperref}
+\hypersetup{
+pdftitle={SIGCHI Conference Proceedings Format},
+pdfauthor={LaTeX},
+pdfkeywords={SIGCHI, proceedings, archival format},
+bookmarksnumbered,
+pdfstartview={FitH},
+colorlinks,
+citecolor=black,
+filecolor=black,
+linkcolor=black,
+urlcolor=black,
+breaklinks=true,
+}
+
+% create a shortcut to typeset table headings
+\newcommand\tabhead[1]{\small\textbf{#1}}
+
 %\linespread{2}
@@
 
-\author{Michael Berg \\ Matthias Untergassmair}
-
-\pagestyle{fancy}
-\rhead{\thepage}
+\numberofauthors{2}
+\author{
+  \alignauthor Matthias Untergassmair\\
+    \email{munter2@uic.edu}\\
+  \alignauthor Michael Berg\\
+    \email{mberg4@uic.edu}\\
+}
 
@@
 
-\abstract{
+\begin{abstract}
 	This paper explores the field of Smooth Particle Hydrodynamics (SPH), starting at its 
@@
 explanation of it.
+\end{abstract}
+
+\keywords{
+	SPH, Smoothed Particle Hydrodynamics, OpenGL, Fluid Simulation, Real Time
 }
 
-\clearpage
+Download code at
+\begin{center}
+	\url{https://github.com/munter2/RealTimeFluid}
+\end{center}
 
@@
 \\
-\hspace*{6 pt} 
+\hspace*{6 pt} The result of Gingold and Monaghan's theory and implementation was a robust 
+and extendable idea that could easily made more accurate "by increasing the number of particles 
+and by using the devices known to improve Monte Carlo integration methods" \cite{sphastrophysics}.
 %
@@
 
+\balance
+
+\nocite{*}
+\printbibliography
+
+\begin{comment}
 \clearpage
@@
 
-\clearpage
-\nocite{*}
-\printbibliography
+\end{comment}
 

### 2014-12-12 Simulation.cpp
Added user interaction + enums for colorcoding and display

@@ -70,15 +70,18 @@ public:
 	Behavior rotateBehavior;
 
 	bool gravityOn;
- 
-
 
 
 	SPH fluidsimulation = SPH(N); // Initialize Fluid simulation model with N particles
 		
 	
 	unsigned stepCounter = 0; // TODO: remove - step counter that keeps track of how many timesteps have been done - model stops after certain number of steps
-	
+
+	enum show { FLUID, FLUIDANDOBJECT, FLUIDANDOBJECTANDWALL };
+	unsigned showParticles = show::FLUID;
+	enum ccodes { METABALLS, NONE, SPEED, VELOCITY, PRESSURE, DENSITY };
+	unsigned colorcoding = ccodes::NONE;
+
 
 
 	void onCreate() {
@@ -88,10 +91,10 @@ public:
 		
 		rh.loadProgram(program, "resources/simulation", posLoc, normalLoc, -1, colLoc);
 
-		M = fluidsimulation.getTotalParticles(); // Render all particles
+		M = fluidsimulation.getTotalParticles();
 		mb = new MeshBuffer[M];
 
-		for(int i=0; i<M; ++i) {
+		for(unsigned i=0; i<M; ++i) {
 			MeshData md;
 			// addCube(md,fluidsimulation.getRadius(i),vec3(0,0,0));
 			// addRect(md,4.f,4.f,100.f,vec3(0,0,0));
@@ -129,6 +132,46 @@ public:
 
 	void onFrame(){
 
+		std::string displayStr;
+		std::string colorcodingStr;
+		
+		switch(colorcoding) {
+			case ccodes::METABALLS:
+				colorcodingStr = "Metaballs";
+				break;
+			case ccodes::NONE:
+				colorcodingStr = "None";
+				break;
+			case ccodes::SPEED:
+				colorcodingStr = "Speed";
+				break;
+			case ccodes::VELOCITY:
+				colorcodingStr = "Velocity";
+				break;
+			case ccodes::PRESSURE:
+				colorcodingStr = "Pressure";
+				break;
+			case ccodes::DENSITY:
+				colorcodingStr = "Density";
+				break;
+		}
+
+		switch(showParticles) {
+			case show::FLUID: // Render fluid only
+				M = fluidsimulation.getFluidParticles();
+				displayStr = "Fluid";
+				break;
+			case show::FLUIDANDOBJECT: // Render fluid and immersed object
[…]

### 2014-12-12 sphModel.cpp

@@ -1,5 +1,11 @@
 #include "sphModel.hpp"
 
+#include <cmath>
+#include <random>
+#include <string>
+
+
+
 // TODO: implement SPH interaction
 // TODO: introduce ghost particles on boundary
 
@@ -7,7 +13,7 @@ SPH::SPH(unsigned N)
 	:	_nParticles(N),
 		_nGhostWall(100),
 		_nGhostObject(0),
-		_nTotal(_nParticles+_nGhostWall+_nGhostObject),
+		_nTotal(_nParticles+_nGhostObject+_nGhostWall),
 		_x1(new float[_nTotal]),
 		_x2(new float[_nTotal]),
 		_x3(new float[_nTotal]),
@@ -19,14 +25,21 @@ SPH::SPH(unsigned N)
 		_a3(new float[_nTotal]),
 		_m(new float[_nTotal]),
 		_r(new float[_nTotal]),
+		_x1MinWall(-100),
+		_x1MaxWall(+100),
 		_x2MinWall(-100),
 		_x2MaxWall(+100),
 		_x3MinWall(-100),
 		_x3MaxWall(+100),
+		_x1MinBox(-40),
+		_x1MaxBox(+40),
 		_x2MinBox(-40),
 		_x2MaxBox(+40),
 		_x3MinBox(-100),
 		_x3MaxBox(+100),
+		_v1Box(0),
+		_v2Box(0),
+		_v3Box(0),
 		_g(0),
 		_damping(.8),
 		_T(0.0),
@@ -71,54 +84,9 @@ SPH::SPH(unsigned N)
 
 	}
 
-	// Initialize Ghost Particles in Wall
-	for(unsigned i=_nParticles; i<_nParticles+_nGhostWall; ++i) {
-		
-		int side = (i-_nParticles)/(.25*_nGhostWall); 
-		float ratio = 0;
-		
-		switch(side) {
-
-			// Position
-			case 0: // Bottom
-				ratio = float(i-_nParticles)/(.25*_nGhostWall); 
-				_x2[i] = -100 + 200*ratio;
-				_x3[i] = -100;
-				break;
-			case 1: // Top
-				ratio = float(i-_nParticles)/(.25*_nGhostWall)-1; 
-				_x2[i] = -100 + 200*ratio;
-				_x3[i] = +100;
-				break;
-			case 2: // Left
-				ratio = float(i-_nParticles)/(.25*_nGhostWall)-2; 
-				_x2[i] = -100;
[…]

### 2014-12-12 sphModel.hpp

@@ -5,8 +5,6 @@
 
 #include <iostream>
 #include <stdio.h>
-#include <cmath>
-#include <random>
 #include <unistd.h>
 
 using std::ostream;
@@ -14,6 +12,8 @@ using std::ostream;
 class SPH {
 
 	public:
+
+		enum _axis { X1, X2, X3 };
 
 		static const unsigned _ghostDepth = 3;
 		
@@ -35,7 +35,9 @@ class SPH {
 		// Applying elastic boundary conditions
 		void applyBoundary();
 
-		// Get Radius of Particle i
+		// Get number of particles
+		unsigned getFluidParticles() const;
+		unsigned getObjectParticles() const;
 		unsigned getTotalParticles() const;
 
 		// Write position to the 3-array x
@@ -75,8 +77,7 @@ class SPH {
 		void setGravity(float);
 
 		// Functions for changing Box position
-		// void moveBoxX(float);
-		// void moveBoxY(float);
+		void moveBox(float,unsigned);
 
 	private:
 
@@ -103,18 +104,23 @@ class SPH {
 		float* _r;
 
 		// Wall Coordinates
+		float _x1MinWall;
+		float _x1MaxWall;
 		float _x2MinWall;
 		float _x2MaxWall;
 		float _x3MinWall;
 		float _x3MaxWall;
 
 		// Box Coordinates
+		float _x1MinBox;
+		float _x1MaxBox;
 		float _x2MinBox;
 		float _x2MaxBox;
 		float _x3MinBox;
 		float _x3MaxBox;
 
 		// Velocity component introduced by Box movement
+		float _v1Box;
 		float _v2Box;
 		float _v3Box;
 

### 2014-12-12 Simulation.cpp
updated README

@@ -53,7 +53,7 @@ class Simulation : public RendererLinux {
 
 public:
 
-	static const unsigned N = 0; // 40;
+	static const unsigned N = 40;
 	unsigned M = 0;
 		
 	ResourceHandler rh; 
@@ -176,9 +176,9 @@ public:
 		// PROPAGATE MODEL
 		///////////////////////////////////////////////////////////
 
-		if(stepCounter < 0 /*5000*/) {
+		if(stepCounter < 5000) {
 			++stepCounter;
-			fluidsimulation.timestep(.05); // Propagate fluidsimulation in time
+			fluidsimulation.timestep(.02); // Propagate fluidsimulation in time
 			std::cout << fluidsimulation; // Output current status of Fluid particles
 		}
 	

### 2014-12-12 sphModel.cpp

@@ -280,11 +280,7 @@ void SPH::updateForces() {
 
 		}
 		_a3[i] += _g; // add gravity
-
 	}
-	std::cout << "Acceleration 1: " << _a1[0];
-	std::cout << "Acceleration 10: " << _a1[9];
-	std::cout << "Acceleration 100: " << _a1[99];
 }
 
 

### 2014-12-12 Simulation.cpp
added colorcoding

@@ -83,7 +83,6 @@ public:
 	unsigned colorcoding = ccodes::NONE;
 
 
-
 	void onCreate() {
 
 		// Output Simulation state
@@ -201,6 +200,8 @@ public:
 		// Draw Cubes
 		float position[3];
 		float velocity[3];
+		vec3 normalizedVelocity;
+		float vmax = fluidsimulation.getVmax();
 		for(unsigned i=0; i<M; ++i) {
 			program.bind(); {
 
@@ -208,6 +209,8 @@ public:
 
 				fluidsimulation.getPosition(i,position);
 				fluidsimulation.getVelocity(i,velocity);
+				normalizedVelocity = vec3(velocity[0],velocity[1],velocity[2]);
+				normalizedVelocity = .5f*(vec3(1.f)+normalizedVelocity/vmax);
 
 				model = glm::translate(model,vec3(position[0],position[1],position[2]));
 
@@ -215,7 +218,8 @@ public:
 				glUniformMatrix4fv(program.uniform("view"), 1, 0, ptr(camera.view));
 				glUniformMatrix4fv(program.uniform("proj"), 1, 0, ptr(camera.projection));
 
-				glUniform3f(program.uniform("velocity"), std::abs(velocity[0])/100,velocity[1]/100,.25);
+				glUniform1i(program.uniform("ccflag"), colorcoding);
+				glUniform3fv(program.uniform("velocity"), 1, ptr(normalizedVelocity));
 
 				mb[i].draw();
 			} program.unbind();

### 2014-12-12 sphModel.cpp

@@ -3,6 +3,7 @@
 #include <cmath>
 #include <random>
 #include <string>
+#include <algorithm>
 
 
 
@@ -213,12 +214,16 @@ void SPH::timestep(float dt) {
 	// Update Forces
 	updateForces();
 
+	_vmax = 0;
+
 	for(unsigned i=0; i<_nParticles; ++i) {
 
 		// Update Velocities
 		_v1[i] += _dt*_a1[i];
 		_v2[i] += _dt*_a2[i];
 		_v3[i] += _dt*_a3[i];
+
+		_vmax = std::max(std::max(_vmax,std::abs(_v1[i])),std::max(std::abs(_v2[i]),std::abs(_v3[i])));
 
 		// Update Positions
 		_x1[i] += _dt*_v1[i];

### 2014-12-12 sphModel.hpp

@@ -69,6 +69,11 @@ class SPH {
 
 		// Return Potential Energy
 		float getEpot() const;
+
+		// Get Maximum Velocity component
+		inline float getVmax() const {
+			return _vmax;
+		}
 		
 		// Get Radius of Particle i
 		float getRadius(unsigned) const;
@@ -123,6 +128,9 @@ class SPH {
 		float _v1Box;
 		float _v2Box;
 		float _v3Box;
+
+		// Keep track of maximum velocity, needed for colorcoding in OpenGL
+		float _vmax; 
 
 
 		// Gravity

### 2014-12-13 Simulation.cpp
Added UIC object

@@ -31,7 +31,6 @@
 
 
 #include "sphModel.hpp"
-#include "extendedShapes.hpp"
 
 
 
@@ -53,7 +52,7 @@ class Simulation : public RendererLinux {
 
 public:
 
-	static const unsigned N = 40;
+	static const unsigned N = 400;
 	unsigned M = 0;
 		
 	ResourceHandler rh; 
@@ -70,14 +69,14 @@ public:
 	Behavior rotateBehavior;
 
 	bool gravityOn;
-
+	bool paused = false;
 
 	SPH fluidsimulation = SPH(N); // Initialize Fluid simulation model with N particles
 		
 	
 
 	enum show { FLUID, FLUIDANDOBJECT, FLUIDANDOBJECTANDWALL };
-	unsigned showParticles = show::FLUID;
+	unsigned showParticles = show::FLUIDANDOBJECTANDWALL;
 	enum ccodes { METABALLS, NONE, SPEED, VELOCITY, PRESSURE, DENSITY };
 	unsigned colorcoding = ccodes::NONE;
 
@@ -94,9 +93,10 @@ public:
 
 		for(unsigned i=0; i<M; ++i) {
 			MeshData md;
-			// addCube(md,fluidsimulation.getRadius(i),vec3(0,0,0));
-			// addRect(md,4.f,4.f,100.f,vec3(0,0,0));
-			addSphere(md,5/*fluidsimulation.getRadius(i)*/,8,8); // TODO: find out what makes it crash here
+			// make sure assertion for r>0 is passed	
+			float r = std::max(fluidsimulation.getRadius(i),.01f);
+			// addCube(md,r);
+			addSphere(md,r,8,8); // TODO: find out what makes it crash here
 			mb[i].init(md,posLoc,normalLoc,-1,colLoc);
 		}
 
@@ -170,7 +170,7 @@ public:
 		}
 
 		// PROPAGATE MODEL AND OUTPUT
-		if(fluidsimulation.getTime() < 10.0) {
+		if(fluidsimulation.getTime() < 10.0 && !paused) {
 		// if(fluidsimulation.getTimeStepNumber() < 10) {
 			fluidsimulation.timestep(.01); // Propagate fluidsimulation in time
 			std::cout << "\nDisplay: " << displayStr << ",\tColorcoding: " << colorcodingStr;
@@ -191,9 +191,16 @@ public:
 		float velocity[3];
 		vec3 normalizedVelocity;
 		float vmax = fluidsimulation.getVmax();
[…]

### 2014-12-13 extendedShapes.cpp

@@ -1,429 +1,0 @@
-#include "Aluminum/Shapes.hpp"
-#include "extendedShapes.hpp"
-
-using glm::vec3;
-
-namespace aluminum {
-
-	void addCube(MeshData &m, float d, vec3 center) {
-
-		float x = center.x;
-		float y = center.y;
-		float z = center.z;
-
-		float s = d/2.f;
-
-		std::cout << "\nPosition Cube:\t(" << x << "," << y << "," << z << ")";
-		
-		//8 vertices
-		vec3 v0 = vec3(-s+x, -s+y, s+z);
-		vec3 v1 = vec3(-s+x, s+y, s+z);
-		vec3 v2 = vec3(s+x, -s+y, s+z);
-		vec3 v3 = vec3(s+x, s+y, s+z);
-		vec3 v4 = vec3(-s+x, -s+y, -s+z);
-		vec3 v5 = vec3(-s+x, s+y, -s+z);
-		vec3 v6 = vec3(s+x, -s+y, -s+z);
-		vec3 v7 = vec3(s+x, s+y, -s+z);
-		
-		//6 normals
-		vec3 n0 = vec3(0, 0, -1.0f);
-		vec3 n1 = vec3(0, 0, +1.0f);
-		vec3 n2 = vec3(0, -1.0f, 0);
-		vec3 n3 = vec3(0, +1.0f, 0);
-		vec3 n4 = vec3(-1.0f, 0, 0);
-		vec3 n5 = vec3(+1.0f, 0, 0);
-		
-		//6 colors
-		/*
-		vec3 c0 = vec3(1.0,0.0,0.0);
-		vec3 c1 = vec3(0.0,1.0,0.0);
-		vec3 c2 = vec3(0.0,0.0,1.0);
-		vec3 c3 = vec3(1.0,1.0,0.0);
-		vec3 c4 = vec3(1.0,0.0,1.0);
-		vec3 c5 = vec3(0.0,1.0,1.0);
-		*/
-
-		vec3 c0 = vec3(.3,.3,.3);
-		vec3 c1 = c0;
-		vec3 c2 = c0;
-		vec3 c3 = c0;
-		vec3 c4 = c0;
-		vec3 c5 = c0;
-		
-		if(x > 0) c0 = vec3(1,0,0);
-		if(x < 0) c1 = vec3(0,1,0);
-		if(y > 0) c2 = vec3(0,0,1);
[…]

### 2014-12-13 extendedShapes.hpp

@@ -1,16 +1,0 @@
-#ifndef EXTENDEDSHAPES_H
-#define EXTENDEDSHAPES_H
-
-#include "Aluminum/MeshData.hpp"
-
-namespace aluminum {
-
-	using glm::vec3;
-
-	void addCube(MeshData&, float, vec3);
-	void addRect(MeshData&, float, float, float, vec3);
-
-}
-
-
-#endif // EXTENDEDSHAPES_H

### 2014-12-13 initObjectCoords.cpp

@@ -1,0 +1,280 @@
+#include "sphModel.hpp"
+
+void SPH::initObjectCoords() {
+	std::cout << "Initializing Object coordinates - make any object here";
+
+
+	// U
+	
+	_uic[0][0] = 0;
+	_uic[1][0] = 0;
+	_uic[2][0] = 1;
+
+	_uic[0][1] = 0;
+	_uic[1][1] = 0;
+	_uic[2][1] = 2;
+	
+	_uic[0][2] = 0;
+	_uic[1][2] = 0;
+	_uic[2][2] = 3;
+
+	_uic[0][3] = 0;
+	_uic[1][3] = 0;
+	_uic[2][3] = 4;
+	
+	_uic[0][4] = 0;
+	_uic[1][4] = 0;
+	_uic[2][4] = 5;
+
+
+
+	_uic[0][5] = 0;
+	_uic[1][5] = 1;
+	_uic[2][5] = 0;
+
+	_uic[0][6] = 0;
+	_uic[1][6] = 1;
+	_uic[2][6] = 1;
+
+	_uic[0][7] = 0;
+	_uic[1][7] = 1;
+	_uic[2][7] = 2;
+
+	_uic[0][8] = 0;
+	_uic[1][8] = 1;
+	_uic[2][8] = 3;
+
+	_uic[0][9] = 0;
+	_uic[1][9] = 1;
+	_uic[2][9] = 4;
+
+	_uic[0][10] = 0;
+	_uic[1][10] = 1;
+	_uic[2][10] = 5;
+
+
+
+	_uic[0][11] = 0;
+	_uic[1][11] = 2;
+	_uic[2][11] = 0;
+
+	_uic[0][12] = 0;
+	_uic[1][12] = 2;
+	_uic[2][12] = 1;
+
+	
+	
+	_uic[0][13] = 0;
+	_uic[1][13] = 3;
+	_uic[2][13] = 0;
+
+	_uic[0][14] = 0;
+	_uic[1][14] = 3;
+	_uic[2][14] = 1;
+
+
+
+	_uic[0][15] = 0;
+	_uic[1][15] = 4;
+	_uic[2][15] = 0;
+
+	_uic[0][16] = 0;
+	_uic[1][16] = 4;
+	_uic[2][16] = 1;
+
+	_uic[0][17] = 0;
+	_uic[1][17] = 4;
+	_uic[2][17] = 2;
+
[…]

### 2014-12-13 sphModel.cpp

@@ -8,10 +8,10 @@
 
 SPH::SPH(unsigned N)
 	:	_nParticles(N),
-		_nGhostObject(10),
+		_nGhostObject(_uicSize),
 		_x1BoxDim(0), // keep it 2d
-		_x2BoxDim(200),
-		_x3BoxDim(200),
+		_x2BoxDim(280),
+		_x3BoxDim(100),
 		_nGhostWall(2*(_x2BoxDim+1+_x3BoxDim+1)*_ghostDepth), // keep it 2d
 		_nTotal(_nParticles+_nGhostObject+_nGhostWall),
 		_x1(new float[_nTotal]),
@@ -40,13 +40,15 @@ SPH::SPH(unsigned N)
 	}
 
 	std::cout << "\nInitializing Model...";
+
+	initObjectCoords();
 	
 	// Seeding random number generator and set parameters for normal distribution
 	// std::random_device rd; // Uncomment to make it even more random ;)
 	// std::mt19937 e2(rd());
 	std::mt19937 e2(42);
 	float mean = 0; // mean velocity
-	float stddev = 10; // standard deviation of velocity
+	float stddev = 4; // standard deviation of velocity
 	std::normal_distribution<> dist(mean,stddev);
 
 	// Initialize Fluid Particles
@@ -76,9 +78,12 @@ SPH::SPH(unsigned N)
 	// Initialize Ghost Particles in Object
 	for(unsigned i=_nParticles; i<(_nParticles+_nGhostObject); ++i) {
 
-		_x1[i] = 0;
-		_x2[i] = 100;
-		_x3[i] = 100;
+		unsigned objectIndex = i - _nParticles;
+
+		_x1[i] = _uic[0][objectIndex];
+		_x2[i] = _uic[1][objectIndex];
+		_x3[i] = _uic[2][objectIndex];
+
 		_v1[i] = 0;
 		_v2[i] = 0;
 		_v3[i] = 0;
@@ -98,28 +103,10 @@ SPH::SPH(unsigned N)
 		
 		zeroIndex = _nParticles + _nGhostObject + 2*d*(_x2BoxDim+1+_x3BoxDim+1);
 
-		std::cout << "\n\n---------------------------------\n d = " << d << "\t\tZEROINDEX = " << zeroIndex;
-
-		for(unsigned i=0; i<(2*_x2BoxDim+1); ++i) {
[…]

### 2014-12-13 sphModel.hpp

@@ -17,7 +17,7 @@ class SPH {
 
 		enum _axis { X1, X2, X3 };
 
-		static const unsigned _ghostDepth = 1;
+		static const unsigned _ghostDepth = 3;
 		
 		// Constructor
 		SPH(unsigned); 
@@ -116,6 +116,11 @@ class SPH {
 		// Functions for changing Box position
 		void moveBox(float,unsigned);
 
+
+		// Create an interesting object
+		void initObjectCoords();
+		
+
 	private:
 
 		unsigned _nParticles; // Number of fluid particles
@@ -161,6 +166,13 @@ class SPH {
 
 		// Was the Box moved?
 		bool _boxMoved;
+
+
+
+
+		// UIC Object
+		static const unsigned _uicSize = 60;
+		float _uic[3][_uicSize];
 		
 };
 

### 2014-12-13 paper/paper.tex

@@
 \\
+\hspace*{6 pt} The set of equations utilized in Gingold and Monaghan's SPH implementation 
+is a density distribution calculated at each point
+\begin{align}
+	\rho_s(r) = \int W(r - r') \rho(r') dr'
+\end{align}
+where W is a function satisfying the condition
+\begin{align}
+\int W(r) dr = 1
+\end{align}
+\hspace*{6 pt} This density distribution equation is used to calculate the density of the 
+particles surrounding the current particle the equation it's being applied to.
+\hspace*{6 pt} A Gaussian function is used for the kernel function, with the form
+\begin{align}
+\exp{\frac{-r^2}{h^2}}
+\end{align}
+where h is
+\begin{align}
+h = b(<r^2> - <r>^2)^\frac{1}{2}
+\end{align}
+and \textit{b} is adjustable.
+\hspace*{6 pt} The gravitational potential of the particles is denoted by
+\begin{align}
+\phi = - G \int \frac{\rho_N(r')}{|r - r'|} dr'
+\end{align}
+\\
 \hspace*{6 pt} The result of Gingold and Monaghan's theory and implementation was a robust 
@@
 and by using the devices known to improve Monte Carlo integration methods" \cite{sphastrophysics}.
+%
+
+\section{Smoothed Particle Hydrodynamics Overview}
+%
+
 %

### 2014-12-13 paper/paper.tex
Finished history section

@@
 
+\clearpage
+
 \section{Introduction}
@@
 \begin{align}
-\exp{\frac{-r^2}{h^2}}
+(\frac{1}{\pi h^2})^\frac{3}{2}  \exp{(\frac{-r^2}{h^2})}
 \end{align}
-where h is
+and h is
 \begin{align}
@@
 \end{align}
+with G being the gravitational constant. Substituting in eq. 1 turns the equation 
+of gravitational potential into
+\begin{align}
+\phi = - \frac{GM}{N} \sum_{j=1}^N \int \frac{W(r - r')}{|r - r'|} dr'
+\end{align}
+To finish the equation, starting with
+\begin{align}
+\triangledown^2I_j = -4 \pi W(r - r_j)
+\end{align}
+substituting gravitational potential for I
+\begin{align}
+\triangledown \rho = - \frac{GM}{N} \sum_{j=1}^N \left \lbrace - \frac{4 \pi}{u_j^2} \int_{0}^{u_j} 
+W(u) u^2 du \right \rbrace \triangledown u_j
+\end{align}
+where
+\begin{align}
+u_j = r - r_j
+\end{align}
+Adding in the Gaussian Function in eq. 3 for W yields
+\begin{align}
+\triangledown \phi = - \frac{GM}{N} \sum_{j=1}^N \frac{2}{u_j} \left ( \frac{f}{\pi} \right )^\frac{1}{2} 
+\left \lbrack \exp{(-f u_j^2)} - \frac{1}{u_j} \int_{0}^{u_j} \exp{(-f u^2)} du \right \rbrack 
+\triangledown u_j
+\end{align}
+This equation is the full solution to calculating the gravitational potential of all particles
+involved in the simulation.
 \\
@@
 
+\clearpage
 \begin{comment}

### 2014-12-14 paper/paper.tex
added cubic spline implementation and derivative

@@
 
+As a kernel function we used the standard kernel, the \emph{cubic spline}
+\begin{align*}
+	W_3(r,h) := \frac{1}{\pi h^3}
+	\begin{cases}
+		1 - \frac{3}{2} \xi^2 + \frac{3}{4} \xi^3, &\text{for } 0 \le \xi < 1 \\
+		\frac{1}{4} (2-\xi)^3, &\text{for } 1 \le \xi < 2 \\
+		0, &\text{otherwise}
+	\end{cases}
+\end{align*}
+%
+where $\xi := \frac{r}{h}$. It can be seen clearly that the support of this function is $2h$. \\
+The derivative is
+%
+\begin{align*}
+	\nabla W_3(r,h) &= \frac{\partial}{\partial r} W_3(r,h) + \frac{\partial}{\partial h} W_3(r,h)
+\end{align*}
+%
+\begin{itemize}
+	\item for $0 \le \xi < 1$
+		\begin{align*}
+			\frac{\partial}{\partial r} W_3(r,h)|_{0 \le \xi < 1}
+			&= \frac{1}{\pi h^3} \frac{1}{h} \left( 3 \xi + \frac{9}{4} \xi^2 \right)
+		\end{align*}
+		%
+		\begin{align*}
+			\frac{\partial}{\partial h} W_3(r,h)|_{0 \le \xi < 1}
+			&= \frac{-3}{\pi h^4} \left( 1 - \frac{3}{2} \xi^2 + \frac{3}{4} \xi^3 \right) \\
+			&\quad + \frac{1}{\pi h^3} \frac{1}{r} \left( (-2) \frac{3}{2} \xi^3 + (-3)\frac{3}{4} \xi^4 \right) \\
+			&= - \frac{3}{\pi h^4} \left( 1 - \frac{3}{2} \xi^2 + \frac{3}{4} \xi^3 \right) \\
+			&\quad + \frac{1}{\pi h^3} \frac{1}{r} \left( -3 \xi^3 -\frac{9}{4} \xi^4 \right)
+		\end{align*}
+		%
+		\begin{align*}
+			\Rightarrow \nabla W_3(r,h)|_{0 \le \xi < 1}
+		&= \frac{1}{\pi h^4} \left( 3 \xi + \frac{9}{4} \xi^2 \right) \\
+			&- \frac{3}{\pi h^4} \left( 1 - \frac{3}{2} \xi^2 + \frac{3}{4} \xi^3 \right) \\
+			&\quad + \frac{1}{\pi h^3} \frac{1}{r} \left( -3 \xi^3 -\frac{9}{4} \xi^4 \right) \\
+		&= \frac{3}{\pi h^4} \left( -1 + \xi + \frac{9}{4} \xi^2
+			- \frac{3}{4} \xi^3 \right) \\
+			&\quad - \frac{1}{\pi r h^3} \xi^3 \left( 3 + \frac{9}{4} \xi \right)
+	\end{align*}
+	\item for $1 \le \xi < 2$
+		\begin{align*}
+			\frac{\partial}{\partial r} W_3(r,h)|_{1 \le \xi < 2}
+			&= \frac{1}{4} 3(2-\xi)^2 \cdot \left( -\frac{1}{h} \right) \\
+			&= -\frac{3}{4h} (2-\xi)^2
+		\end{align*}
+		%
+		\begin{align*}
+			\frac{\partial}{\partial h} W_3(r,h)|_{1 \le \xi < 2}
+			&= \frac{1}{4} 3(2-\xi)^2 \cdot \left( +\frac{\xi}{h} \right) \\
+			&= \frac{3 \xi}{4 h} (2-\xi)^2
+		\end{align*}
+		\begin{align*}
+			\Rightarrow \nabla W_3(r,h)|_{1 \le \xi < 2}
+			&= -\frac{3}{4h} (2-\xi)^2
+			+ \frac{3 \xi}{4 h} (2-\xi)^2 \\
+			&= \frac{3}{4h} (2-\xi)^2 (\xi - 1)
+		\end{align*}
+	\item otherwise
+		\begin{align*}
+			\nabla W_3(r,h)|_{\text{otherwise}} = 0
+		\end{align*}
+\end{itemize}
+%
+The following two code snippets show how these formulas are implemented in our simulation:
+\begin{lstlisting}[caption=Cubic Spline]
+// Cubic spline function
+inline float W3(float r, float h) {
+	float W = 0;
+	float zeta = r/h;
+	if(0 <= zeta && zeta <= 2) {
+		W = (zeta < 1 ? 
+				1 - .75*zeta*zeta*(2+zeta) :
+				.25*(2-zeta)*(2-zeta)*(2-zeta)
+		);
+	}
+	W /= (M_PI*h*h*h);
+	return W;
+}
+\end{lstlisting}
+%
+\begin{lstlisting}[caption=Derivative of Cubic Spline]
+Hello World
+\end{lstlisting}
+
 \end{document}

### 2014-12-14 Simulation.cpp

@@ -52,7 +52,7 @@ class Simulation : public RendererLinux {
 
 public:
 
-	static const unsigned N = 400;
+	static const unsigned N = 40;
 	unsigned M = 0;
 		
 	ResourceHandler rh; 

### 2014-12-14 sphModel.cpp

@@ -14,6 +14,7 @@ SPH::SPH(unsigned N)
 		_x3BoxDim(100),
 		_nGhostWall(2*(_x2BoxDim+1+_x3BoxDim+1)*_ghostDepth), // keep it 2d
 		_nTotal(_nParticles+_nGhostObject+_nGhostWall),
+		_nActive(0),
 		_x1(new float[_nTotal]),
 		_x2(new float[_nTotal]),
 		_x3(new float[_nTotal]),
@@ -28,6 +29,8 @@ SPH::SPH(unsigned N)
 		_vmax(1e-10),
 		_g(0),
 		_damping(.8),
+		_support(10),
+		_h(.5*_support),
 		_T(0.0),
 		_tStep(0),
 		_dt(0),
@@ -47,17 +50,17 @@ SPH::SPH(unsigned N)
 	// std::random_device rd; // Uncomment to make it even more random ;)
 	// std::mt19937 e2(rd());
 	std::mt19937 e2(42);
-	float mean = 0; // mean velocity
-	float stddev = 4; // standard deviation of velocity
+	float mean = 50; // mean velocity
+	float stddev = 10; // standard deviation of velocity
 	std::normal_distribution<> dist(mean,stddev);
 
 	// Initialize Fluid Particles
 	for(unsigned i=0; i<_nParticles; ++i) {
 		
-		// Position
+		// Position (Locate at source)
 		_x1[i] = 0; // keep it 2d
-		_x2[i] = fmod(rand(),_x2BoxDim)-.5*_x2BoxDim;
-		_x3[i] = fmod(rand(),_x3BoxDim)-.5*_x3BoxDim;
+		_x2[i] = -.45*_x2BoxDim; // fmod(rand(),_x2BoxDim)-.5*_x2BoxDim;
+		_x3[i] = .45*_x3BoxDim; // fmod(rand(),_x3BoxDim)-.5*_x3BoxDim;
 		
 		// Masses (assume two groups of particle masses)
 		_m[i] = .5*(1+i%2);
@@ -71,7 +74,7 @@ SPH::SPH(unsigned N)
 		// Velocities (sampled from random normal distribution)
 		_v1[i] = 0; // keep it 2d
 		_v2[i] = dist(e2);
-		_v3[i] = dist(e2);
+		_v3[i] = 0; // dist(e2);
 
 	}
 
@@ -97,7 +100,6 @@ SPH::SPH(unsigned N)
 
[…]

### 2014-12-14 sphModel.hpp

@@ -54,6 +54,34 @@ class SPH {
 			return _tStep;
 		}
 
+		// Cubic spline function
+		inline float W3(float r, float h) {
+			float W = 0;
+			float xi = r/h;
+			if(0 <= xi && xi <= 2) {
+				W = (xi < 1 ? 
+						1 - .75*xi*xi*(2+xi) :
+						.25*(2-xi)*(2-xi)*(2-xi)
+				);
+			}
+			W /= (M_PI*h*h*h);
+			return W;
+		}
+
+		// Cubic spline derivative
+		inline float deltaW3(float r, float h) {
+			float W = 0;
+			float xi = r/h;
+			if(0 <= xi && xi <= 2) {
+				W = (xi < 1 ? 
+						1/(h*h*h*M_PI) * (.75/h*(-4+4*xi+9*xi*xi-3*xi*xi*xi) + .25/r*xi*xi*xi*(12+9*xi)) :
+						.75/h*(2-xi)*(2-xi)*(xi-1)		
+				);
+			}
+			W /= (M_PI*h*h*h);
+			return W;
+		}
+
 
 
 		// Write position to the 3-array x
@@ -130,6 +158,7 @@ class SPH {
 		unsigned _x3BoxDim;
 		unsigned _nGhostWall; // Number of ghost particles in the walls
 		unsigned _nTotal; // Total number of particles
+		unsigned _nActive;
 
 		// Array of particle coordinates & velocities & accelerations
 		float* _x1;
@@ -156,6 +185,8 @@ class SPH {
 
 		// Damping factor for elastic bounding on walls
 		float _damping;
+		float _support;
+		float _h;
 
 		// Total time
 		float _T;

### 2014-12-14 paper/paper.tex
Added some SPH stuff to article

@@
 \usepackage{listings}
+\usepackage{placeins}
 
@@
 %
-As suggested in \cite{generalizedwallboundary}, we use the \emph{Velocity-Verlet} time stepping scheme as follows:
-%
-\begin{algorithm}
-	\caption{Single Timestep with Velocity Verlet Algorithm}
-	\KwData{$x_i^t$, $v_i^{t - \frac{\Delta t}{2}}$, $a_i^t$, $\Delta t$}
-	\KwResult{$x_i^{t + \Delta t}$, $v_i^{t + \frac{\Delta t}{2}}$, $a_i^{t + \Delta t}$}	
-	$v_i^{t+\frac{\Delta t}{2}} = v_i^{t - \frac{\Delta t}{2}} + \Delta t a_i^t$ \;
-	$x_i^{t+\Delta t} = x_i^t + \Delta t v_i^{t+\frac{\Delta t}{2}}$ \;	
-	$a_i^{t + \Delta t} = a_i^{t + \Delta t} (x_i^{t + \Delta t}, m_i)$ from equation \ref{math:acceleration} \;
-\end{algorithm}
 
@@
 \end{comment}
+
+\clearpage
+
+Several resources point out the algorithms to be used for applying SPH to Fluid Dynamics, the algorithm is described particularly clearly in \cite{kennyerleben} and can be summarized as follows:
+%
+\FloatBarrier
+\begin{algorithm}
+	\caption{SPH simulation}
+	\For{$i = 1, \dots, N_{\text{Particles}}$}{
+		Compute the density of the $i$th particle
+		\begin{align*}
+			\rho_i = \sum_j m_j W(\| x_i - x_j \|, h)
+		\end{align*}
+		Compute the pressure of the $i$th particle
+		\begin{align*}
+			p_i = k (\rho_i - \rho_0)
+		\end{align*}
+		Compute the interaction forces acting on the $i$th particle, consisting of pressure forces, viscosity and gravity
+		\begin{align*}
+			f_i^p &= - \sum_j m_j \frac{p_i + p_j}{2 \rho_j} \nabla W(\| x_i - x_j \|, h) \\
+			f_i^v &= \mu \sum_j m_j \frac{v_j + v_i}{\rho_j} \nabla^2 W(\| x_i - x_j \|, h) \\
+			f_i^{\text{total}} &= f_i^p + f_i^v + g
+		\end{align*}
+		Solve the differential equation
+		\begin{align}
+			\begin{aligned}
+				\dot{x}_i &= v_i \\
+				\dot{v}_i &= \frac{f_i}{\rho_i}
+			\end{aligned}
+			\label{math:ode}
+		\end{align}
+	}
+\end{algorithm}
+\FloatBarrier
+%
+In our implementation, we use the Velocity Verlet (or Leap-Frog) scheme to solve the ODE \ref{math:ode}, as suggested in \cite{generalizedwallboundary}.
+The scheme looks as follows:
+%
+\begin{algorithm}
+	\caption{Single Timestep with Velocity Verlet Algorithm}
+	\KwData{$x_i^t$, $v_i^{t - \frac{\Delta t}{2}}$, $a_i^t$, $\Delta t$}
+	\KwResult{$x_i^{t + \Delta t}$, $v_i^{t + \frac{\Delta t}{2}}$, $a_i^{t + \Delta t}$}	
+	$v_i^{t+\frac{\Delta t}{2}} = v_i^{t - \frac{\Delta t}{2}} + \Delta t a_i^t$ \;
+	$x_i^{t+\Delta t} = x_i^t + \Delta t v_i^{t+\frac{\Delta t}{2}}$ \;	
+	$a_i^{t + \Delta t} = a_i^{t + \Delta t} (x_i^{t + \Delta t}, m_i)$ from equation \ref{math:acceleration} \;
+\end{algorithm}
+
+\clearpage
 

### 2014-12-14 sphModel.cpp

@@ -25,12 +25,15 @@ SPH::SPH(unsigned N)
 		_a2(new float[_nTotal]),
 		_a3(new float[_nTotal]),
 		_m(new float[_nTotal]),
+		_rho(new float[_nTotal]),
+		_p(new float[_nTotal]),
 		_r(new float[_nTotal]),
 		_vmax(1e-10),
 		_g(0),
 		_damping(.8),
 		_support(10),
 		_h(.5*_support),
+		_kPressure(1),
 		_T(0.0),
 		_tStep(0),
 		_dt(0),
@@ -64,6 +67,8 @@ SPH::SPH(unsigned N)
 		
 		// Masses (assume two groups of particle masses)
 		_m[i] = .5*(1+i%2);
+		_rho[i] = 1;
+		_p[i] = 1;
 
 		// Radius / Support of particles
 		_r[i] = 1; // 1+i%3;
@@ -93,6 +98,7 @@ SPH::SPH(unsigned N)
 	
 		_r[i] = 1;
 		_m[i] = 1;
+		_p[i] = 1;
 
 	}
 	
@@ -116,6 +122,7 @@ SPH::SPH(unsigned N)
 			_v3[zeroIndex+i] = 0;
 
 			_m[zeroIndex+i] = 1e10;
+			_p[zeroIndex+i] = 1;
 			_r[zeroIndex+i] = .4;
 		}
 
@@ -132,6 +139,7 @@ SPH::SPH(unsigned N)
 			_v3[zeroIndex+i] = 0;
 
 			_m[zeroIndex+i] = 1e10;
+			_p[zeroIndex+i] = 1;
 			_r[zeroIndex+i] = .4;
 		}
 	
@@ -141,6 +149,8 @@ SPH::SPH(unsigned N)
 SPH::~SPH() {
 	// Free memory
 	if(_r) { delete[] _r; }
+	if(_p) { delete[] _p; }
+	if(_rho) { delete[] _rho; }
 	if(_m) { delete[] _m; }
 	if(_a3) { delete[] _a3; }
 	if(_a2) { delete[] _a2; }
@@ -224,7 +234,7 @@ void SPH::updateForces() {
 
 			// Compute Force
 			F = 1000*W3(R,_h);
-			// F = (R<10 && R>.001 ? -10/(R*R*R) : 0); // .1*(R>10 ? R : -1000000/R); // TODO: replace by SPH force computation
[…]

### 2014-12-14 sphModel.hpp

@@ -173,6 +173,12 @@ class SPH {
 
 		// Array of particle masses
 		float* _m;
+
+		// Array of particle densities
+		float* _rho;
+		
+		// Array of particle pressures
+		float* _p;
 		
 		// Array of particle radii
 		float* _r;
@@ -187,6 +193,7 @@ class SPH {
 		float _damping;
 		float _support;
 		float _h;
+		float _kPressure;
 
 		// Total time
 		float _T;

### 2014-12-14 paper/paper.tex
Added Overview section

--- 2014-12-14 paper/paper.tex
@@
 %
-
+\hspace{6 pt} Smoothed-particle hydrodynamics (SPH) method works by dividing the fluid into a set of 
+referred to as particles. These particles have a spatial distance, or smoothing length, over which 
+their properties are "smoothed" by a kernel function, such as a Gaussian function
+\begin{align}
+(\frac{1}{\pi h^2})^\frac{3}{2}  \exp{(\frac{-r^2}{h^2})}
+\end{align}
+This means that the physical quantity of any particle can be obtained by summing the 
+relevant properties of all the particles which lie within the range of the kernel. For 
+example, using Monaghan's popular cubic spline kernel the temperature at position r 
+depends on the temperatures of all the particles within a radial distance 2h of r \cite{wikiarticle}.
+\\
+\hspace*{6 pt} The contributions of each particle to a property are weighted according to their distance 
+from the particle of interest, and their density
+A property for any one particle is calculated by discerning how many particles lay within
+smoothing length of a given particle and what their respective properties are, and how 
+dense that smoothing length is around the particle of interest. Mathematically, this is 
+governed by a kernel function, like the Gaussian function given in eq. 1. This Gaussian
+function gives large numbers for particles whose smoothing lengths are close to the 
+particle of interest, and thus effect the calculation of a given particles properties, 
+but as particles get far away, and their smoothing length becomes extremely small, 
+although never reaches 0. This is due to intrinsic properties of the 
+Gaussian function itself, but it is easy to account for this problem by testing for 
+extremely small values close to zero, and disregarding those particles that are far away 
+from the particle of interest as a result.
+\\
+\hspace*{6 pt} The equation for any quantity A at any point r is given by the equation
+\begin{align}
+A(r) = \sum_{j} m_j \frac{a_j}{p_j} W (|r - r_j|,h)
+\end{align}
+where m is the mass of particle j, A is the value of the quantity A for particle j, r 
+denotes the position, p is the particle's density and W is a kernel function, such as the 
+Gaussian function used by Gingold and Monaghan shown in eq. 1. The density of 
+the particle can be computed in a variety of ways, such as the example equation
+\begin{align}
+p_i = p(r_i) = \sum_{j} m_j \frac{p_j}{p_j} W (|r_i - r_i|,h) = \sum_{j} m_j W(r_i - r_j,h)
+\end{align}
+Similarly, the spatial derivative of a quantity can be computed as follows
+\begin{align}
+\triangledown A(r) = \sum_{j} m_j \frac{a}{p_j} \triangledown W(|r - r_j|, h)
+\end{align}
+Although the size of the smoothing length can be fixed in both space and time, this does 
+not take advantage of the full power of SPH. By assigning each particle its own smoothing 
+length and allowing it to vary with time, the resolution of a simulation can be made to 
+automatically adapt itself depending on local conditions. For example, in a very dense 
+region where many particles are close together the smoothing length can be made relatively 
+short, yielding high spatial resolution. Conversely, in low-density regions where 
+individual particles are far apart and the resolution is low, the smoothing length can be 
+increased, optimising the computation for the regions of interest. Combined with an 
+equation of state and an integrator, SPH can simulate hydrodynamic flows efficiently \cite{wikiarticle}.
+\\
+\hspace*{6 pt} Given the principles of smoothing length, calculating properties based on 
+the area around a particle, and increasing or decreasing a property given the density of 
+particles in an area, we applied some of these concepts in our implementation of SPH.
 %

### 2014-12-14 Simulation.cpp
beautiful

@@ -43,8 +43,6 @@ float pi = glm::pi<float>();
 using namespace aluminum;
 
 
-// TODO: improve performance by only adding one single sphere instead of N spheres (just use different model matrices).
-// TODO: make liquid flow in from top
 // TODO: pass in only points to shader and use geometry shader to create 3d particles
 // TODO: see 3.5.1: flowing water and particle effects, stream output
 
@@ -52,7 +50,7 @@ class Simulation : public RendererLinux {
 
 public:
 
-	static const unsigned N = 40;
+	static const unsigned N = 4;
 	unsigned M = 0;
 		
 	ResourceHandler rh; 
@@ -171,7 +169,7 @@ public:
 
 		// PROPAGATE MODEL AND OUTPUT
 		if(fluidsimulation.getTime() < 10.0 && !paused) {
-		// if(fluidsimulation.getTimeStepNumber() < 10) {
+		// if(fluidsimulation.getTimeStepNumber() < 10 && !paused) {
 			fluidsimulation.timestep(.01); // Propagate fluidsimulation in time
 			std::cout << "\nDisplay: " << displayStr << ",\tColorcoding: " << colorcodingStr;
 			std::cout << fluidsimulation; // Output current status of Fluid particles

### 2014-12-14 sphModel.cpp

@@ -34,6 +34,8 @@ SPH::SPH(unsigned N)
 		_support(10),
 		_h(.5*_support),
 		_kPressure(1),
+		_rho0(1),
+		_mu(0),
 		_T(0.0),
 		_tStep(0),
 		_dt(0),
@@ -121,7 +123,7 @@ SPH::SPH(unsigned N)
 			_v2[zeroIndex+i] = 0;
 			_v3[zeroIndex+i] = 0;
 
-			_m[zeroIndex+i] = 1e10;
+			_m[zeroIndex+i] = 1;
 			_p[zeroIndex+i] = 1;
 			_r[zeroIndex+i] = .4;
 		}
@@ -138,7 +140,7 @@ SPH::SPH(unsigned N)
 			_v2[zeroIndex+i] = 0;
 			_v3[zeroIndex+i] = 0;
 
-			_m[zeroIndex+i] = 1e10;
+			_m[zeroIndex+i] = 1;
 			_p[zeroIndex+i] = 1;
 			_r[zeroIndex+i] = .4;
 		}
@@ -175,7 +177,8 @@ void SPH::timestep(float dt) {
 	_T += _dt;
 	++_tStep;
 
-	// Update Forces
+	// Update Parameters
+	updateDensityPressure();
 	updateForces();
 
 	_vmax = 0;
@@ -203,26 +206,60 @@ void SPH::timestep(float dt) {
 
 }
 
-void SPH::updateForces() {
+void SPH::updateDensityPressure() {
 
 	float d1, d2, d3; // Particle Distance in each space direction
 	float R; // Particle Distance in 3D space
-	float theta, phi; // Angles for orientation in 3D space
-	float F; // Force between two particles
 
-	for(unsigned i=0; i<_nParticles; ++i) { // TODO: put nActive
-		
-		_a1[i] = 0;
-		_a2[i] = 0;
-		_a3[i] = 0;
+	// Compute density and prassure
+	for(unsigned i=0; i<_nTotal; ++i) {
+
+		_rho[i] = 0;
+		_p[i] = 0;
 
 		for(unsigned a=0; a<_nTotal; ++a) {
-
+			
 			if(a == i) continue; // Particles don't interact with themselves
 			
 			d1 = _x1[a] - _x1[i];
 			d2 = _x2[a] - _x2[i];
 			d3 = _x3[a] - _x3[i];
+
+			R = sqrt(d1*d1+d2*d2+d3*d3);
+
[…]

### 2014-12-14 sphModel.hpp

@@ -30,6 +30,9 @@ class SPH {
 		
 		// Update forces on particles based on SPH
 		void updateForces();
+		
+		// Update density and pressure based on SPH
+		void updateDensityPressure();
 
 		// Overloading Output Operator
 		friend ostream& operator<<(ostream&, const SPH&);
@@ -69,7 +72,21 @@ class SPH {
 		}
 
 		// Cubic spline derivative
-		inline float deltaW3(float r, float h) {
+		inline float d1W3(float r, float h) {
+			float W = 0;
+			float xi = r/h;
+			if(0 <= xi && xi <= 2) {
+				W = (xi < 1 ? 
+						1/(h*h*h*M_PI) * (.75/h*(-4+4*xi+9*xi*xi-3*xi*xi*xi) + .25/r*xi*xi*xi*(12+9*xi)) :
+						.75/h*(2-xi)*(2-xi)*(xi-1)		
+				);
+			}
+			W /= (M_PI*h*h*h);
+			return W;
+		}
+		
+		// Cubic spline second derivative
+		inline float d2W3(float r, float h) {
 			float W = 0;
 			float xi = r/h;
 			if(0 <= xi && xi <= 2) {
@@ -194,6 +211,8 @@ class SPH {
 		float _support;
 		float _h;
 		float _kPressure;
+		float _rho0; // Environment pressure
+		float _mu; // Viscosity
 
 		// Total time
 		float _T;

### 2014-12-14 Simulation.cpp
handin version

@@ -50,7 +50,7 @@ class Simulation : public RendererLinux {
 
 public:
 
-	static const unsigned N = 4;
+	static const unsigned N = 400;
 	unsigned M = 0;
 		
 	ResourceHandler rh; 
@@ -94,7 +94,7 @@ public:
 			// make sure assertion for r>0 is passed	
 			float r = std::max(fluidsimulation.getRadius(i),.01f);
 			// addCube(md,r);
-			addSphere(md,r,8,8); // TODO: find out what makes it crash here
+			addSphere(md,r,8,8);
 			mb[i].init(md,posLoc,normalLoc,-1,colLoc);
 		}
 
@@ -168,7 +168,8 @@ public:
 		}
 
 		// PROPAGATE MODEL AND OUTPUT
-		if(fluidsimulation.getTime() < 10.0 && !paused) {
+		if(!paused) {
+		// if(fluidsimulation.getTime() < 10.0 && !paused) {
 		// if(fluidsimulation.getTimeStepNumber() < 10 && !paused) {
 			fluidsimulation.timestep(.01); // Propagate fluidsimulation in time
 			std::cout << "\nDisplay: " << displayStr << ",\tColorcoding: " << colorcodingStr;
@@ -283,7 +284,7 @@ public:
 				fluidsimulation.setGravity(0);
 				gravityOn = false;
 			} else {
-				fluidsimulation.setGravity(-20);
+				fluidsimulation.setGravity(-200);
 				gravityOn = true;
 			}
 		

### 2014-12-14 sphModel.cpp

@@ -123,7 +123,7 @@ SPH::SPH(unsigned N)
 			_v2[zeroIndex+i] = 0;
 			_v3[zeroIndex+i] = 0;
 
-			_m[zeroIndex+i] = 1;
+			_m[zeroIndex+i] = 1e3;
 			_p[zeroIndex+i] = 1;
 			_r[zeroIndex+i] = .4;
 		}
@@ -217,13 +217,14 @@ void SPH::updateDensityPressure() {
 		_rho[i] = 0;
 		_p[i] = 0;
 
-		for(unsigned a=0; a<_nTotal; ++a) {
+		for(unsigned j=0; j<_nTotal; ++j) {
+		
+			if(j < _nParticles && j > _nActive) continue; // Don't interact with inactive particles
+			if(j == i) continue; // Particles don't interact with themselves
 			
-			if(a == i) continue; // Particles don't interact with themselves
-			
-			d1 = _x1[a] - _x1[i];
-			d2 = _x2[a] - _x2[i];
-			d3 = _x3[a] - _x3[i];
+			d1 = _x1[j] - _x1[i];
+			d2 = _x2[j] - _x2[i];
+			d3 = _x3[j] - _x3[i];
 
 			R = sqrt(d1*d1+d2*d2+d3*d3);
 
@@ -255,6 +256,7 @@ void SPH::updateForces() {
 
 		for(unsigned j=0; j<_nTotal; ++j) {
 			
+			if(j < _nParticles && j > _nActive) continue; // Don't interact with inactive particles
 			if(j == i) continue; // Particles don't interact with themselves
 			
 			d1 = _x1[j] - _x1[i];
@@ -270,8 +272,8 @@ void SPH::updateForces() {
 			theta = acos(d3/R);
 
 			// Compute Force
-			Fp = (_rho[j] != 0 ? _m[j]*(_p[i]+_p[j])/(2*_rho[j]) * d1W3(R,_h) : 0); // TODO: uncomment
-			Fv = 0; // _mu*_m[j]*(_v[j]+_v[i])/_rho[j] * d2W3(R,_h); // TODO: uncomment
+			Fp = (_rho[j] != 0 ? _m[j]*(_p[i]+_p[j])/(2*_rho[j]) * d1W3(R,_h) : 0);
+			Fv = 0; // _mu*_m[j]*(_v[j]+_v[i])/_rho[j] * d2W3(R,_h); // TODO: implement d2w3
 			F = Fp + Fv;			
 			F *= 100;
 
[…]

### 2014-12-14 sphModel.hpp

@@ -37,9 +37,6 @@ class SPH {
 		// Overloading Output Operator
 		friend ostream& operator<<(ostream&, const SPH&);
 
-		// Applying elastic boundary conditions
-		void applyBoundary();
-
 		// Get number of particles
 		inline unsigned getFluidParticles() const {
 			return _nParticles;

### 2014-12-14 paper/paper.tex
beautiful
Video Added

@@
 
+\setlength\parindent{12pt}
+
 %\linespread{2}
@@
 implementation as a way to simulate fluids such as water. We will then give a brief 
-mathematical background and ensuing algorithm of our SPH simulation followed by a detailed 
-explanation of it.
+mathematical background and ensuing algorithm of our SPH simulation followed by a brief explanation of the code and an outlook for possible improvements.
 \end{abstract}
@@
 
-\clearpage
-
 \section{Introduction}
 %
-\hspace{6 pt} Smoothed Particle Hydrodynamics (SPH) successfully simulates fluids by breaking up a 
+Smoothed Particle Hydrodynamics (SPH) successfully simulates fluids by breaking up a 
 fluid body into individual parts, or particles. These particles together form a particle 
-system that simulates various gravitational forces. Fluid movement is simulated in the 
-system by moving particles around any particle moved, simulating a rippling, wave effect.
-\\
-\hspace*{6 pt} The ripple effect is created by first calculating which particles are surrounding 
-a particle moved, and these surrounding particles are moved according to the movement 
-of the first particle moved. But SPH wasn't originally intended to simulate liquid 
-substances, but astrophysical phenomena.
+system that simulates physical properties of the system based on various interaction forces and external forces. 
+The governing equations for fluid movement can be deduced from the governing \emph{Navier Stokes} equations.
+\par\medskip
+However, SPH wasn't originally intended to be used in the simulation of liquid substances, but for investigating astrophysical phenomena.
 %
@@
 %
-\hspace{6 pt} Given all the different applications for Smoothed Particle Hydrodynamics (SPH),
+Given all the different applications for Smoothed Particle Hydrodynamics (SPH),
 it was first used to simulate interstellar phenomena. Conceived in 1977 by Gingold and 
@@
 "move according to the Newtonian equations with forces due to the pressure gradient and 
-other body forces: gravity, rotation and magnetic" \cite{sphastrophysics}.
-\\
-\hspace*{6 pt} The set of equations utilized in Gingold and Monaghan's SPH implementation 
+other body forces: gravity, rotation and magnetic" \cite{sphastrophysics}. \\
+From the Lagrangian perspective, the computational elements are not fixed in space, but they move around as particles and carry several physical properties with them.
+\par\medskip
+The set of equations utilized in Gingold and Monaghan's SPH implementation 
 is a density distribution calculated at each point
@@
 \end{align}
-where W is a function satisfying the condition
+where $W$ is a so called \emph{kernel function} satisfying the normalization condition
 \begin{align}
 	\int W(r) dr = 1
+	\label{math:kernelnormal}
 \end{align}
-\hspace*{6 pt} This density distribution equation is used to calculate the density of the 
+%
+\par\medskip
+This density distribution equation is used to calculate the density of the 
 particles surrounding the current particle the equation it's being applied to.
-\hspace*{6 pt} A Gaussian function is used for the kernel function, with the form
+\par\medskip
+%
+An example for a kernel function is the Gaussian
 \begin{align}
-(\frac{1}{\pi h^2})^\frac{3}{2}  \exp{(\frac{-r^2}{h^2})}
+ \left( \frac{1}{\pi h^2} \right)^\frac{3}{2}  \exp{(\frac{-r^2}{h^2})}
 \end{align}
-and h is
+where $h$ is
 \begin{align}
-h = b(<r^2> - <r>^2)^\frac{1}{2}
+h = b( \langle r^2 \rangle - \langle r \rangle^2)^\frac{1}{2}
 \end{align}
-and \textit{b} is adjustable.
-\hspace*{6 pt} The gravitational potential of the particles is denoted by
+and $b$ is an adjustable parameter.
+\par\medskip
+%
+The gravitational potential of the particles is denoted by
 \begin{align}
@@
 involved in the simulation.
-\\
-\hspace*{6 pt} The result of Gingold and Monaghan's theory and implementation was a robust 
+\par\medskip
+%
+The result of Gingold and Monaghan's theory and implementation was a robust 
 and extendable idea that could easily made more accurate "by increasing the number of particles 
 and by using the devices known to improve Monte Carlo integration methods" \cite{sphastrophysics}.
+\par\medskip
 %
+We would like to point out here the similarity of the kernel function $W$ to the \emph{Dirac Delta} function
+\begin{align*}
+	\delta(x) := 
+	\begin{cases}
+		+ \infty, & x =0 \\
+		0 , & x \neq 0
+	\end{cases}
+\end{align*}
+which is also constrained to satisify the normalization condition
+\begin{align*}
+	\int_{-\infty}^\infty \delta (x) dx = 1
+\end{align*}
+The similarity to the Gaussian function becomes apparent when looking at the limit
+\begin{align*}
+	&\delta_a (x) \xrightarrow{a \rightarrow 0} \delta(x)
+	&\delta_a (x) := \frac{1}{a \sqrt{\pi}} \exp \left\{ \frac{-x^2}{a^2} \right\}
+\end{align*}
 
@@
 %
-\hspace{6 pt} Smoothed-particle hydrodynamics (SPH) method works by dividing the fluid into a set of 
+Smoothed-particle hydrodynamics (SPH) method works by dividing the fluid into a set of 
 referred to as particles. These particles have a spatial distance, or smoothing length, over which 
@@
 \begin{align}
-(\frac{1}{\pi h^2})^\frac{3}{2}  \exp{(\frac{-r^2}{h^2})}
+\left( \frac{1}{\pi h^2} \right)^\frac{3}{2}  \exp{\left( \frac{-r^2}{h^2} \right)}
 \end{align}
-This means that the physical quantity of any particle can be obtained by summing the 
+This means that the physical quantity of any particle can be obtained by approximating the continuous integral \ref{math:kernelnormal} and summing the 
 relevant properties of all the particles which lie within the range of the kernel. For 
-example, using Monaghan's popular cubic spline kernel the temperature at position r 
+example, using Monaghan's popular cubic spline kernel the temperature at position $r$ 
 depends on the temperatures of all the particles within a radial distance 2h of r \cite{wikiarticle}.
-\\
-\hspace*{6 pt} The contributions of each particle to a property are weighted according to their distance 
+\par\medskip
+%
+The contributions of each particle to a property are weighted according to their distance 
 from the particle of interest, and their density
@@
 from the particle of interest as a result.
-\\
-\hspace*{6 pt} The equation for any quantity A at any point r is given by the equation
+\par\medskip
+%
+In general, the equation for any quantity $A$ at any point $r$ is given by the equation
 \begin{align}
-A(r) = \sum_{j} m_j \frac{a_j}{p_j} W (|r - r_j|,h)
+A(r) = \sum_{j} m_j \frac{a_j}{\rho_j} W (|r - r_j|,h)
 \end{align}
-where m is the mass of particle j, A is the value of the quantity A for particle j, r 
-denotes the position, p is the particle's density and W is a kernel function, such as the 
+where $m$ is the mass of particle $j$, $A$ is the value of the quantity $A$ for particle $j$, $r$ 
+denotes the position, $\rho$ is the particle's density and $W$ is a kernel function, such as the 
 Gaussian function used by Gingold and Monaghan shown in eq. 1. The density of 
@@
 individual particles are far apart and the resolution is low, the smoothing length can be 
-increased, optimising the computation for the regions of interest. Combined with an 
+increasee, optimising the computation for the regions of interest. Combined with an 
 equation of state and an integrator, SPH can simulate hydrodynamic flows efficiently \cite{wikiarticle}.
-\\
-\hspace*{6 pt} Given the principles of smoothing length, calculating properties based on 
+\par\medskip
+%
+Given the principles of smoothing length, calculating properties based on 
 the area around a particle, and increasing or decreasing a property given the density of 
@@
 %
-
-\section{Mathematical Background}
-%
-\begin{align}
-	a_i^n = \frac{F_i^n}{m_i} = \dots
-	\label{math:acceleration}
-\end{align}
+\clearpage
 
@@
 We omit the vector notation ($\boldsymbol{x},\boldsymbol{v},\boldsymbol{a}$) for these quantities, since the following equations are valid for the vectors as well as for each component individually.
-\par\medskip
 %
-
-\balance
-
-\nocite{*}
-\printbibliography
-
-\clearpage
-\begin{comment}
-\clearpage
-
-\section*{Appendix: Code}
-
-\lstinputlisting[caption=sphModel.hpp]{../sphModel.hpp}
-\lstinputlisting[caption=sphModel.cpp]{../sphModel.cpp}
-\lstinputlisting[caption=Simulation.cpp]{../Simulation.cpp}
-\lstinputlisting[caption=simulation.vsh]{../resources/simulation.vsh}
-\lstinputlisting[caption=simulation.fsh]{../resources/simulation.fsh}
-
-\end{comment}
-
-\clearpage
-
 Several resources point out the algorithms to be used for applying SPH to Fluid Dynamics, the algorithm is described particularly clearly in \cite{kennyerleben} and can be summarized as follows:
@@
 \end{algorithm}
-
-\clearpage
-
+\par\medskip
+%
 As a kernel function we used the standard kernel, the \emph{cubic spline}
@@
 The following two code snippets show how these formulas are implemented in our simulation:
+\FloatBarrier
 \begin{lstlisting}[caption=Cubic Spline]
@@
 	float W = 0;
-	float zeta = r/h;
-	if(0 <= zeta && zeta <= 2) {
-		W = (zeta < 1 ? 
-				1 - .75*zeta*zeta*(2+zeta) :
-				.25*(2-zeta)*(2-zeta)*(2-zeta)
+	float xi = r/h;
+	if(0 <= xi && xi <= 2) {
+		W = (xi < 1 ? 
+				1 - .75*xi*xi*(2+xi) :
+				.25*(2-xi)*(2-xi)*(2-xi)
 		);
@@
 \end{lstlisting}
+\FloatBarrier
 %
-\begin{lstlisting}[caption=Derivative of Cubic Spline]
+\begin{lstlisting}[caption=First Derivative of Cubic Spline]
+// Cubic spline derivative
+inline float d1W3(float r, float h) {
+	float W = 0;
+	float xi = r/h;
+	if(0 <= xi && xi <= 2) {
+		W = (xi < 1 ? 
+				1/(h*h*h*M_PI) * (.75/h*(-4+4*xi+9*xi*xi-3*xi*xi*xi) + .25/r*xi*xi*xi*(12+9*xi)) :
+				.75/h*(2-xi)*(2-xi)*(xi-1)		
+		);
+	}
+	W /= (M_PI*h*h*h);
+	return W;
+}
+\end{lstlisting}
+\FloatBarrier
+%
+\begin{comment}
+\begin{lstlisting}[caption=Second Derivative of Cubic Spline]
 Hello World
 \end{lstlisting}
+\end{comment}
+%
+
+\section{Possible Improvements}
+%
+Our Implementation of SPH is not yet suitable for being used as a real time fluid rendering tool.
+The code would be easily extendible to three dimensions, since the simulation already now makes use of all three space dimensions. \\
+However, the code scales pretty badly with the number of particles:
+The number of particles grows exponentially with the number of dimensions and the complexity of the algorithm grows quadratically with the number of particles!
+More sophisticated algorithms must be applied in order to tackle this bottleneck.
+\par\medskip
+%
+Possible improvements for our simulation include
+\begin{itemize}
+	\item Introducing smart data structures that allow to reduce the complexity when looking for nearest neighbors (for example cell lists, spatial hashing, etc.).
+	\item Render the Fluid in an appealing way using metaballs
+	\item 
+\end{itemize}<++>
+
+\balance
+
+\nocite{*}
+\printbibliography
+
+\clearpage
+\begin{comment}
+\clearpage
+
+\section*{Appendix: Code}
+
+\lstinputlisting[caption=sphModel.hpp]{../sphModel.hpp}
+\lstinputlisting[caption=sphModel.cpp]{../sphModel.cpp}
+\lstinputlisting[caption=Simulation.cpp]{../Simulation.cpp}
+\lstinputlisting[caption=simulation.vsh]{../resources/simulation.vsh}
+\lstinputlisting[caption=simulation.fsh]{../resources/simulation.fsh}
+
+\end{comment}
+
+\clearpage
+
+\clearpage
 
